\section{Methodology}
\label{sec:method}

We formalize the adaptive question selection problem and describe the 14 algorithms evaluated in our study.

\subsection{Problem Formulation}
\label{sec:formulation}

\paragraph{Student model.}
Each student maintains a knowledge vector $\mathbf{k} = [k_1, \ldots, k_K] \in [0,1]^K$ over $K$ topics.
When quizzed on topic $i$, the student answers correctly with probability $k_i$ (Bernoulli trial).
Knowledge evolves according to two dynamics:

\noindent\textit{Learning.}
After answering a question on topic $i$:
\begin{equation}
    k_i \leftarrow
    \begin{cases}
        k_i + \alpha \cdot (1 - k_i), & \text{if correct}, \\
        k_i - \beta \cdot k_i, & \text{if incorrect},
    \end{cases}
    \label{eq:learning}
\end{equation}
where $\alpha = 0.12$ is the learning rate and $\beta = 0.02$ is the incorrect-answer penalty.
The asymptotic form ensures diminishing returns as knowledge approaches mastery.

\noindent\textit{Forgetting.}
At each timestep, all topics \emph{not} quizzed undergo Ebbinghaus-inspired exponential decay~\citep{ebbinghaus1885memory}:
\begin{equation}
    k_j \leftarrow k_{\mathrm{base}} + (k_j - k_{\mathrm{base}}) \cdot e^{-\lambda}, \quad \forall j \neq i,
    \label{eq:forgetting}
\end{equation}
where $k_{\mathrm{base}} = 0.10$ is the retention floor and $\lambda > 0$ is the per-step decay rate.
Higher $\lambda$ corresponds to faster forgetting.

\paragraph{RMAB formulation.}
We cast this problem as a restless multi-armed bandit~\citep{whittle1988restless}.
Each of the $K$ topics is an arm.
The state of arm~$i$ at time~$t$ is its knowledge level $k_i^{(t)} \in [0,1]$.
At each timestep the learner selects one arm to activate (quiz).
The activated arm transitions according to \eqref{eq:learning}, while all passive arms decay according to \eqref{eq:forgetting}.
The objective is to maximize cumulative mean knowledge:
\begin{equation}
    \max_{\pi} \;\; \E\!\left[\frac{1}{T} \sum_{t=1}^{T} \bar{k}^{(t)}\right], \quad \text{where } \bar{k}^{(t)} = \frac{1}{K}\sum_{i=1}^{K} k_i^{(t)}.
    \label{eq:objective}
\end{equation}
This RMAB is intractable in general (PSPACE-hard~\citep{papadimitriou1999complexity}), motivating the heuristic algorithms below.

\subsection{Baseline Algorithms}
\label{sec:baselines}

We evaluate seven baseline algorithms spanning classical bandits, heuristics, and a myopic greedy policy.

\paragraph{Random.}
Selects a topic uniformly at random at each step.
Serves as the primary baseline; all improvements are reported relative to Random.

\paragraph{UCB1~\citep{auer2002finite}.}
Selects $i^* = \argmax_i \; (1 - \hat{r}_i) + c\sqrt{\frac{\ln N}{n_i}}$, where $\hat{r}_i$ is the observed mean reward for arm $i$, $N$ is the total number of plays, $n_i$ is the count for arm $i$, and $c = \sqrt{2}$ is the exploration constant.
The inverted reward $(1 - \hat{r}_i)$ targets weak topics.

\paragraph{Thompson Sampling~\citep{thompson1933likelihood}.}
Maintains a Beta posterior $\text{Beta}(\alpha_i, \beta_i)$ for each arm.
At each step, samples $\theta_i \sim \text{Beta}(\alpha_i, \beta_i)$ and selects $i^* = \argmax_i (1 - \theta_i)$.

\paragraph{$\varepsilon$-Greedy.}
With probability $\varepsilon = 0.1$ selects a random arm; otherwise selects the arm with the lowest observed mean reward.

\paragraph{Sliding-Window UCB (SW-UCB)~\citep{garivier2011upper}.}
Restricts reward estimates to the most recent $\tau$ observations per arm, adapting to non-stationarity.
Uses $\tau = 50$ in our experiments.

\paragraph{Leitner System~\citep{leitner1972so}.}
Implements the classical box-based spaced repetition heuristic.
Topics start in box~1 and advance on correct answers; incorrect answers reset to box~1.
Lower boxes are reviewed more frequently.

\paragraph{Greedy-Min.}
Has access to the true knowledge vector $\mathbf{k}^{(t)}$ at each step and selects $i^* = \argmin_i \; k_i^{(t)}$---the topic with the lowest current knowledge.%
\footnote{Despite having access to the true knowledge state, Greedy-Min is myopic---it does not plan ahead or account for future forgetting dynamics. It should not be confused with a dynamic-programming-optimal policy.}
Despite perfect information, this myopic greedy policy does not account for future forgetting.

\subsection{Novel Algorithms}
\label{sec:novel}

We introduce five algorithms specifically designed for the forgetting-aware selection problem.

\paragraph{Forgetting-UCB (F-UCB).}
F-UCB augments the standard UCB score with a decay-aware urgency term.
Let $t_i$ be the number of steps since arm $i$ was last played and $\hat{r}_i$ the observed mean reward.
The score is:
\begin{equation}
    s_i^{\text{F-UCB}} = \underbrace{(1 - \hat{r}_i) \cdot e^{-\lambda t_i}}_{\text{weighted weakness}} + \underbrace{\gamma \cdot (1 - e^{-\lambda t_i})}_{\text{decay urgency}} + \underbrace{c\sqrt{\frac{\ln N}{n_i}}}_{\text{exploration}},
    \label{eq:fucb}
\end{equation}
where $\gamma = 1.0$ controls the urgency weight and $\lambda$ is the known decay rate.
The first term downweights the observed weakness estimate as it becomes stale; the second term increases urgency for arms that have been decaying longer; the third is the standard UCB exploration bonus.

\paragraph{F-UCB (Decay).}
F-UCB (Decay) is an ablation of F-UCB that retains only the decay-urgency terms, omitting the UCB exploration bonus:
\begin{equation}
    s_i^{\text{F-UCB(D)}} = (1 - \hat{r}_i) \cdot e^{-\lambda t_i} + \gamma \cdot (1 - e^{-\lambda t_i}).
    \label{eq:fucb_decay}
\end{equation}
Comparing F-UCB (Decay) to full F-UCB isolates the contribution of the exploration bonus in the forgetting setting.

\paragraph{BKT-Bandit.}
BKT-Bandit maintains a Bayesian posterior over each arm's mastery probability, augmented with forgetting.
Each arm $i$ has parameters $(\alpha_i, \beta_i)$ of a Beta distribution, updated on observations:
\begin{equation}
    \alpha_i \leftarrow \alpha_i + \mathbb{1}[\text{correct}], \quad
    \beta_i \leftarrow \beta_i + \mathbb{1}[\text{incorrect}].
\end{equation}
Between observations, the posterior is decayed toward uncertainty:
\begin{equation}
    \alpha_i \leftarrow 1 + (\alpha_i - 1) \cdot e^{-\lambda}, \quad
    \beta_i \leftarrow 1 + (\beta_i - 1) \cdot e^{-\lambda}.
\end{equation}
The exponential posterior shrinkage toward the prior can be interpreted as a discounted posterior~\citep{raj2017taming}, where the effective sample size of past observations decays geometrically with rate $e^{-\lambda}$. This is equivalent to a Bayesian update under a hazard model where the student's knowledge state may change (due to forgetting) with probability $1 - e^{-\lambda}$ per timestep, justifying the use of a time-varying posterior.

The selection score combines weakness with uncertainty:
\begin{equation}
    s_i^{\text{BKT}} = \underbrace{(1 - \mu_i)}_{\text{weakness}} + \underbrace{c \cdot \sigma_i}_{\text{uncertainty bonus}},
    \label{eq:bkt}
\end{equation}
where $\mu_i = \frac{\alpha_i}{\alpha_i + \beta_i}$ and $\sigma_i = \sqrt{\frac{\alpha_i \beta_i}{(\alpha_i + \beta_i)^2(\alpha_i + \beta_i + 1)}}$ are the posterior mean and standard deviation.

\paragraph{Advantage Index.}
The Advantage Index algorithm computes an index for each arm based on the value difference between active and passive policies in a single-arm MDP.
For arm $i$ with current state $k_i$, the advantage index approximates:
\begin{equation}
    W_i = V_i^{\text{active}}(k_i) - V_i^{\text{passive}}(k_i),
    \label{eq:whittle}
\end{equation}
where $V^{\text{active}}$ is the value of quizzing the arm (applying \eqref{eq:learning}) and $V^{\text{passive}}$ is the value of leaving it idle (applying \eqref{eq:forgetting}).
We solve the single-arm MDP via value iteration on a discretized state space ($M{=}50$ bins).
The arm with the highest advantage index is selected at each step.
Unlike classical Whittle indices~\citep{whittle1988restless}, which require a Lagrangian relaxation and verified indexability, our advantage index is a heuristic value-difference that does not assume indexability.
Empirically, the single-arm value function exhibits threshold structure (high advantage for low-knowledge states, low for high-knowledge states), suggesting approximate indexability holds in practice.
We tested posterior-based and adaptive variants with negligible performance differences; we report the base version.

\paragraph{MetaSelector.}
MetaSelector maintains a portfolio of $M$ base selectors (BKT-Bandit, F-UCB, Advantage Index, Leitner, and Random in our experiments).
At each step, each base selector proposes a candidate category.
MetaSelector then scores each unique candidate using a unified formula that combines F-UCB's three-term structure with BKT-Bandit's Bayesian posterior:
\begin{equation}
    s_i^{\text{meta}} = \underbrace{(1 - \mu_i) \cdot e^{-\lambda t_i}}_{\text{decayed weakness}} + \underbrace{0.5 \cdot (1 - e^{-\lambda t_i})}_{\text{urgency}} + \underbrace{c \cdot \sigma_i}_{\text{uncertainty}},
    \label{eq:meta}
\end{equation}
where $\mu_i$ and $\sigma_i$ are the posterior mean and standard deviation from a Beta posterior with forgetting (as in BKT-Bandit), and $t_i$ is the time since arm $i$ was last played.
The category with the highest score is selected.
This design connects to Optimistic Online Mirror Descent~\citep{orabona2025online}: the urgency term acts as a ``hint'' about future losses from decay, while the expert candidates restrict the action set from $K$ to at most $M$ alternatives, reducing effective regret~\citep{orabona2025online}.

\paragraph{EquiBandit.}
EquiBandit extends BKT-Bandit with an exposure-equity penalty to address the equity--efficiency tradeoff observed in our experiments (\Cref{sec:regime}).
The score penalizes categories that have received disproportionately many quizzes:
\begin{equation}
    s_i^{\text{equi}} = (1 - \mu_i) + c \cdot \sigma_i - \eta \cdot \max\!\left(0,\; \frac{n_i}{\bar{n}} - 1\right),
\end{equation}
where $n_i$ is the number of times category $i$ has been quizzed, $\bar{n}$ is the average across categories, and $\eta = 0.1$ is the equity weight.
This regularizer discourages over-concentration on a subset of categories without abandoning weakness-targeting entirely.

\subsection{Theoretical Analysis}
\label{sec:theory}

We provide theoretical grounding for the key empirical findings.

\begin{theorem}[Greedy-Min Catastrophe Condition]
\label{thm:catastrophe}
Consider $K$ arms with identical initial knowledge $k_0$, learning rate $\alpha$, penalty $\beta$, decay rate $\lambda$, and retention floor $k_{\mathrm{base}}$. The myopic greedy policy $\pi_{\mathrm{greedy}} = \argmin_i k_i^{(t)}$ is dominated by balanced allocation whenever
\begin{equation}
    K > 1 + \frac{(\alpha - \beta)\, k_0 (1 - k_0)}{\lambda\,(k_0 - k_{\mathrm{base}})}.
    \label{eq:catastrophe_condition}
\end{equation}
Under this condition, the cumulative regret of the greedy policy relative to round-robin grows as $\Omega(K \lambda T)$.
\end{theorem}

\begin{proof}[Proof sketch]
At each timestep, the greedy policy quizzes one arm while $K{-}1$ arms decay. The per-step change in total knowledge $S = \sum_i k_i$ under greedy is $\Delta S_{\mathrm{greedy}} = (\alpha-\beta)k(1-k) - (K-1)\lambda(k-k_{\mathrm{base}}) + O(\lambda^2)$. When condition~\eqref{eq:catastrophe_condition} holds, $\Delta S_{\mathrm{greedy}} < 0$: total knowledge decreases each step. This drives greedy toward a low-knowledge steady state $\bar{k}_{\mathrm{greedy}}^*$. Meanwhile, the balanced (round-robin) policy converges to a higher steady state $\bar{k}_{\mathrm{balanced}}^* > \bar{k}_{\mathrm{greedy}}^*$ because each arm is quizzed once per $K$-step cycle and the cycle-gain can offset decay. The steady-state gap satisfies $\bar{k}_{\mathrm{balanced}}^* - \bar{k}_{\mathrm{greedy}}^* = \Omega(\lambda)$, yielding cumulative regret $\Omega(K\lambda T)$ (Supplementary~\ref{app:proofs}).
\end{proof}

For our default parameters ($\alpha{=}0.12$, $\beta{=}0.02$, $k_0{=}0.5$, $k_{\mathrm{base}}{=}0.10$), the catastrophe threshold is $K > 1 + 0.025/(\lambda \cdot 0.4) = 1 + 0.0625/\lambda$, yielding $K > 13.5$ at $\lambda{=}0.005$ and $K > 2.25$ at $\lambda{=}0.05$---consistent with our experimental observations (\Cref{sec:oracle}).

\begin{theorem}[BKT-Bandit Posterior Concentration]
\label{thm:bkt}
Under BKT-Bandit's posterior decay mechanism with rate $e^{-\lambda}$, the effective sample size for arm $i$ after $n_i$ observations converges to $n_{\mathrm{eff}} = \min\!\bigl(n_i,\; \frac{1}{1 - e^{-\lambda}}\bigr)$. The posterior mean $\mu_i$ concentrates around the true knowledge $k_i^*$ with high probability:
\begin{equation}
    \Pr\bigl[|\mu_i - k_i^*| > \varepsilon\bigr] \leq 2\exp\!\bigl(-2\, n_{\mathrm{eff}}\, \varepsilon^2\bigr).
    \label{eq:bkt_concentration}
\end{equation}
This bound is conservative (valid): the geometric decay makes $n_{\mathrm{eff}}$ a lower bound on information content, so the right-hand side is an upper bound on the actual concentration probability.
Consequently, BKT-Bandit achieves Bayesian regret $\mathcal{O}(K\sqrt{T \log T})$.
\end{theorem}

\begin{proof}[Proof sketch]
The geometric decay $\alpha_i \leftarrow 1 + (\alpha_i - 1) e^{-\lambda}$ acts as a discount on past observations, yielding effective counts $\alpha_i^{\mathrm{eff}} = 1 + \sum_{s \leq t} e^{-\lambda(t-s)} \cdot \mathbb{1}[\text{correct at } s]$. The geometric sum saturates at $1/(1{-}e^{-\lambda}) \approx 1/\lambda$ for small $\lambda$. Standard Beta posterior concentration~\citep{lattimore2020bandit} applies with $n$ replaced by $n_{\mathrm{eff}}$. The Bayesian regret bound follows from the posterior sampling analysis of~\citet{russo2014learning}, adapted to the effective sample size regime (Supplementary~\ref{app:proofs}).
\end{proof}

\begin{theorem}[F-UCB Sublinear Regret]
\label{thm:fucb}
F-UCB with known decay rate $\lambda$ achieves cumulative regret $\mathcal{O}(\sqrt{KT\log T})$ relative to the best fixed-arm policy, provided $\lambda > 0$.
\end{theorem}

\begin{proof}[Proof sketch]
The exponential decay $e^{-\lambda t_i}$ in the F-UCB score~\eqref{eq:fucb} discounts stale estimates and bounds effective regime changes per arm to $\mathcal{O}(1/\lambda)$. Within each regime, standard UCB analysis~\citep{auer2002finite} gives $\mathcal{O}(\log T/\Delta_i)$ suboptimal plays per arm. Evaluating at the minimax-optimal gap scale $\Delta_i = \Theta(\sqrt{K\log T / T})$, the per-arm regret is $O(\log T / \Delta_i) = O(\sqrt{T\log T / K})$, and summing over $K$ arms yields $K \cdot O(\sqrt{T\log T/K}) = O(\sqrt{KT\log T})$. The urgency term $\gamma(1-e^{-\lambda t_i})$ prevents arm starvation, guaranteeing $\Omega(T/K)$ plays per arm and avoiding the Greedy-Min catastrophe (\Cref{thm:catastrophe}). Full proof in Supplementary~\ref{app:proofs}.
\end{proof}

\begin{theorem}[Minimax Lower Bound]
\label{thm:lower_bound}
Any algorithm $\pi$ for the $K$-armed forgetting bandit with horizon $T$ and decay rate $\lambda \geq 0$ incurs expected cumulative regret relative to the best fixed-arm policy:
\begin{equation}
    R_T(\pi) \;\geq\; \Omega\!\left(\sqrt{KT}\right).
    \label{eq:lower_bound}
\end{equation}
Consequently, F-UCB is minimax-optimal up to $\mathcal{O}(\sqrt{\log T})$ factors.
\end{theorem}

\begin{proof}[Proof sketch]
Any forgetting-bandit instance with $\lambda \to 0$ reduces exactly to a $K$-armed stochastic bandit (no arm transitions when not played). By Theorem~15.2 of~\citet{lattimore2020bandit}, the minimax regret for $K$-armed bandits is $\Omega(\sqrt{KT})$. Since the forgetting bandit strictly generalizes the standard bandit, the lower bound is inherited.
Together with \Cref{thm:fucb}, F-UCB achieves $\mathcal{O}(\sqrt{KT\log T})$, matching this lower bound up to $\sqrt{\log T}$. Full proof in Supplementary~\ref{app:proofs}.
\end{proof}

\begin{remark}
In the high-decay regime ($\lambda \gg \sqrt{K/T}$), forgetting dominates exploration: each step where the optimal arm is idle costs $\lambda(k^* - k_{\mathrm{base}})$ in recovered knowledge, creating an additional penalty proportional to $\lambda T$ for algorithms forced to explore. This regime is precisely where F-UCB's urgency term $\gamma(1-e^{-\lambda t_i})$ provides the largest empirical gains (Theorem~\ref{thm:catastrophe}, \Cref{sec:regime}).
\end{remark}
