\setcounter{page}{1}

\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thesection}{S\arabic{section}}
\setcounter{section}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

\begin{center}
\Large\textbf{Forgetting-Aware Bandits for Adaptive Learning:\\From UCB to Restless Experts}\par
\vspace{1em}
\Large Supplementary Material
\end{center}

\noindent In this supplementary document, we provide:
\begin{itemize}
    \item \textbf{Section~\ref{app:details}:} Experimental details and compute resources.
    \item \textbf{Section~\ref{app:additional}:} Additional results and consistency analysis.
    \item \textbf{Section~\ref{app:pseudocode}:} Algorithm pseudocode for MetaSelector.
    \item \textbf{Section~\ref{app:realdata}:} Real data fitting details.
    \item \textbf{Section~\ref{app:whittle}:} Whittle index implementation details.
    \item \textbf{Section~\ref{app:limitations}:} Extended limitations discussion.
\end{itemize}

% ============================================================
\section{Experimental Details}
\label{app:details}

\paragraph{Student model parameters.}
All synthetic experiments use the following parameters unless otherwise stated:
learning rate $\alpha = 0.12$,
incorrect-answer penalty $\beta = 0.02$,
decay rates $\lambda \in \{0.005, 0.01, 0.05\}$,
retention floor $k_{\mathrm{base}} = 0.10$,
UCB exploration constant $c = \sqrt{2} \approx 1.414$.
Each configuration uses 100 students, each answering 10{,}000 questions.
Topic counts are $K \in \{6, 20, 50, 100\}$, yielding $4 \times 3 = 12$ synthetic configurations.
Each configuration is repeated with 3 random seeds; we report mean $\pm$ standard error.

\paragraph{Hardware and runtime.}
All experiments were run on the KAUST IBEX high-performance computing cluster using the \texttt{batch} partition.
Each job was allocated 4 CPU cores and 32~GB RAM; no GPUs were used.
Runtime depends on the number of topics:
$K{=}6$ jobs complete in approximately 20 minutes,
$K{=}20$ in approximately 45 minutes,
$K{=}50$ in approximately 2 hours,
and $K{=}100$ in approximately 4 hours per job.
The total compute budget for all 12 configurations $\times$ 14 algorithms $\times$ 3 seeds = 504 jobs was approximately 800 CPU-hours.

\paragraph{Software.}
Experiments were implemented in Python 3.11 using NumPy ($\geq$1.24), SciPy ($\geq$1.11), Matplotlib ($\geq$3.7), Seaborn ($\geq$0.12), and Pandas ($\geq$2.0).
The simulation environment follows the Gymnasium~\citep{towers2024gymnasium} interface.
Conda environment: \texttt{chessgcn} on IBEX.

\paragraph{Initialization.}
Initial knowledge levels are sampled from a truncated normal distribution:
$k_i^{(0)} \sim \mathrm{clip}(\mathcal{N}(\mu \cdot d_i, \sigma^2), 0.05, 0.95)$
where $\mu = 0.35$, $\sigma = 0.1$, and $d_i \sim \mathrm{Uniform}(0.3, 0.7)$.
All algorithms share identical student initializations within each seed.

% ============================================================
\section{Additional Results}
\label{app:additional}

\paragraph{Consistency analysis.}
\Cref{tab:consistency} reports the consistency of each algorithm across all 12 synthetic configurations.
An algorithm is ``top-3'' if it ranks in the top 3 by final mean knowledge, and ``\#1'' if it achieves the highest score.
Average rank is computed across all 12 configurations.

\begin{table}[h]
\centering
\caption{Algorithm consistency across 12 synthetic configurations ($K \in \{6,20,50,100\}$, $\lambda \in \{0.005, 0.01, 0.05\}$).}
\label{tab:consistency}
\input{tables/table_consistency}
\end{table}

BKT-Bandit achieves the best average rank (2.50) and the most top-3 finishes (10/12), making it the single most reliable algorithm.
F-UCB ranks second with 4 wins but is less consistent across regimes.
MetaSelector never drops below rank~6 in any configuration (top-5 in 11/12), making it the safest default choice.

\paragraph{Temporal dynamics.}
Knowledge trajectories reveal qualitatively different phases.
In the first 1{,}000 steps, exploration-heavy algorithms (UCB1, Thompson) lag behind exploitation-heavy ones (Greedy-Min, $\varepsilon$-Greedy).
Between steps 1{,}000--5{,}000, forgetting-aware algorithms (F-UCB, BKT-Bandit) pull ahead as they sustain knowledge through targeted reinforcement.
Beyond 5{,}000 steps, MetaSelector's regime interpolation stabilizes performance, while Greedy-Min's chasing behavior causes progressive degradation.

\paragraph{Area Under Knowledge Curve (AUC).}
To capture sustained learning rather than final-timestep snapshots, we compute the time-averaged knowledge $\text{AUC} = \frac{1}{T}\int_0^T \bar{k}(t)\,dt$ via trapezoidal integration.
At $K{=}6$, $\lambda{=}0.01$, PD-Whittle achieves the highest AUC (0.575), followed by Greedy-Min (0.575) and Whittle (0.574).
At $K{=}20$, $\lambda{=}0.01$, BKT-Bandit leads with AUC 0.193, followed by F-UCB (0.186).
AUC rankings largely mirror final-knowledge rankings but reveal that Whittle variants sustain knowledge more consistently in low-decay regimes.

\paragraph{Time-to-mastery.}
We measure the first timestep where $\min_i k_i(t) \geq 0.50$ (i.e., all categories reach 50\% mastery).
At $K{=}6$, $\lambda{=}0.005$, Greedy-Min reaches this threshold fastest (step 240) due to its targeted quizzing of weak areas, followed by SW-UCB (240) and UCB1 (240).
At $K{=}6$, $\lambda{=}0.01$, only four algorithms ever reach 50\% mastery: Greedy-Min (step 457), SW-UCB (497), UCB1 (523), and Whittle (547).
For $\lambda{=}0.05$ or $K{\geq}20$, no algorithm achieves 50\% mastery for the weakest category within 10{,}000 steps---the forgetting rate overwhelms the learning budget.

\paragraph{Equity--efficiency trade-off.}
Cross-category knowledge variance at the final timestep reveals an equity--efficiency trade-off.
Greedy-Min achieves near-zero variance ($<10^{-5}$) but poor average knowledge.
BKT-Bandit, the strongest average-knowledge algorithm at $K{=}20$, has the \emph{worst} variance---it concentrates on high-value categories at the expense of weaker ones.
This trade-off suggests that combining BKT-Bandit with equity-aware methods could yield both high average knowledge and balanced learning.

\paragraph{Exposure equity (Gini coefficient).}
We measure the Gini coefficient of the per-category exposure distribution at the final timestep.
A Gini of 0 indicates perfectly equal exposure; higher values indicate concentration on fewer categories.
At $K{=}6$, $\lambda{=}0.01$, all algorithms achieve near-perfect equity (Gini $<0.007$).
At $K{=}20$, $\lambda{=}0.01$, most algorithms remain equitable (Gini $<0.03$), but BKT-Bandit concentrates heavily on a subset of categories (Gini $= 0.449$, ``very unequal''), confirming its equity--efficiency trade-off.
MetaSelector exhibits moderate concentration (Gini $= 0.107$), while Whittle variants and Random maintain near-perfect equity (Gini $<0.003$).

\paragraph{F-UCB urgency weight ($\gamma$) sensitivity.}
We vary the urgency weight $\gamma \in \{0.0, 0.1, 0.3, 0.5, 0.7, 1.0\}$ at $K{=}6$.
At $\lambda{=}0.01$, performance increases monotonically with $\gamma$: from $\bar{k}{=}0.298$ ($\gamma{=}0$) to $0.587$ ($\gamma{=}1.0$), confirming that urgency is critical.
At $\lambda{=}0.05$, the relationship is non-monotone: performance peaks at $\gamma{=}0.1$ ($\bar{k}{=}0.241$) and degrades for larger $\gamma$ ($\bar{k}{=}0.143$ at $\gamma{=}1.0$), suggesting that excessive urgency causes thrashing at high decay.
The default $\gamma{=}0.5$ is a reasonable compromise, but regime-specific tuning could improve results by 10--20\%.

% ============================================================
\section{Algorithm Pseudocode}
\label{app:pseudocode}

\begin{algorithm}[h]
\caption{MetaSelector: UCB over Expert Selectors}
\label{alg:meta}
\begin{algorithmic}[1]
\REQUIRE Base selectors $\mathcal{E} = \{e_1, \ldots, e_M\}$, exploration constant $c_{\mathrm{meta}}$, decay rate $\lambda$
\STATE Initialize counts $N_j \leftarrow 0$, rewards $R_j \leftarrow 0$ for each expert $j$
\FOR{$t = 1, 2, \ldots, T$}
    \IF{any expert has $N_j = 0$}
        \STATE $j^* \leftarrow$ any unexplored expert \COMMENT{Round-robin init}
    \ELSE
        \STATE $j^* \leftarrow \argmax_j \frac{R_j}{N_j} + c_{\mathrm{meta}} \sqrt{\frac{\ln \sum_m N_m}{N_j}}$ \COMMENT{UCB over experts}
    \ENDIF
    \STATE $i^* \leftarrow e_{j^*}.\textsc{SelectArm}(\text{history})$ \COMMENT{Expert picks topic}
    \STATE Play arm $i^*$, observe reward $r_t$
    \STATE Update student knowledge via Eqs.~(\ref{eq:learning})--(\ref{eq:forgetting})
    \STATE $\bar{k}_t \leftarrow \frac{1}{K}\sum_{i=1}^{K} k_i^{(t)}$ \COMMENT{Current mean knowledge}
    \STATE $N_{j^*} \leftarrow N_{j^*} + 1$
    \STATE $R_{j^*} \leftarrow R_{j^*} + \bar{k}_t$ \COMMENT{Credit mean knowledge to expert}
\ENDFOR
\end{algorithmic}
\end{algorithm}

The unified scoring function used within MetaSelector for its internal ranking is:
\begin{equation}
    s_i^{\text{unified}} = (1 - \mu_i) \cdot e^{-\lambda t_i} + 0.5 \cdot (1 - e^{-\lambda t_i}) + c \cdot \sigma_i,
    \tag{S1}
\end{equation}
where $\mu_i$ and $\sigma_i$ are the posterior mean and standard deviation for arm $i$, $t_i$ is the time since last play, and $\lambda$ is the decay rate.

% ============================================================
\section{Real Data Details}
\label{app:realdata}

\paragraph{Duolingo.}
We use the Duolingo English learner dataset~\citep{settles2016trainable}, sampling 500 students with at least 50 interactions, yielding 127{,}000 total interactions.
Knowledge topics correspond to lexeme tags.
We fit student model parameters via maximum likelihood estimation on held-out data:
\begin{itemize}[leftmargin=*,itemsep=1pt]
    \item Learning rate: $\alpha = 0.126$
    \item Incorrect penalty: $\beta = 0.087$
    \item Decay rate: $\lambda \approx 0$ (effectively zero)
    \item Initial knowledge: $k_0 = 0.91$
\end{itemize}
The near-zero decay rate is consistent with Duolingo's built-in spaced repetition system already mitigating forgetting.
In this regime, most algorithms perform similarly, and the Greedy-Min catastrophe does not manifest strongly.

\paragraph{ASSISTments.}
We use the ASSISTments 2009--2010 skill-builder dataset~\citep{feng2009addressing}, sampling 500 students with at least 100 interactions, yielding 425{,}000 total interactions.
Knowledge topics correspond to skill tags.
Fitted parameters:
\begin{itemize}[leftmargin=*,itemsep=1pt]
    \item Learning rate: $\alpha = 0.090$
    \item Incorrect penalty: $\beta = 0.077$
    \item Decay rate: $\lambda = 0.008$
    \item Initial knowledge: $k_0 = 0.71$
\end{itemize}
The non-trivial decay ($\lambda = 0.008$) makes ASSISTments a more informative testbed.
BKT-Bandit and PD-Whittle maintain their advantage, and the Greedy-Min catastrophe is visible (Greedy-Min underperforms Random).

% ============================================================
\section{Whittle Index Implementation Details}
\label{app:whittle}

The Whittle advantage index is computed on a discretized state space with $M=50$ bins over $k \in [0.01, 0.99]$. Value iteration uses convergence threshold $10^{-8}$ with maximum 500 iterations. The advantage index $W(k) = V_{\text{mixed}}(k) - V_{\text{passive}}(k)$ uses mixed transition probability $1/K$ to account for the budget constraint. Precomputation takes ${\sim}0.01$ seconds per configuration at $M{=}50$; online lookup is $O(1)$ via nearest-state interpolation.

\paragraph{Sensitivity to discretization.}
We tested $M \in \{20, 50, 100, 200\}$ and found that coarser grids ($M{=}20, 50$) produce stable, monotonically decreasing index profiles---high index for low-mastery states, as expected. Finer grids ($M{=}100, 200$) introduce numerical instabilities in the binary-search/value-iteration pipeline, producing non-monotone index profiles. We therefore use $M{=}50$ as the default, which balances approximation quality with numerical stability. All granularities compute in under 0.1 seconds.

% ============================================================
\section{Extended Limitations}
\label{app:limitations}

\begin{itemize}[leftmargin=*,itemsep=2pt]
    \item \textbf{Synthetic--real gap.} Our synthetic environment uses a simplified student model with homogeneous learning/forgetting parameters across topics.
    Real students exhibit topic-dependent learning rates, prerequisite dependencies, and motivational effects not captured in our simulation.
    The near-zero decay fitted on Duolingo suggests that real platforms' existing scheduling may already mitigate the forgetting that drives our synthetic results.

    \item \textbf{Specific failure modes.} BKT-Bandit underperforms at $K{=}20$, $\lambda{=}0.005$, likely because its posterior decay mechanism introduces unnecessary uncertainty in near-stationary settings.
    MetaSelector's single win ($K{=}6$, $\lambda{=}0.01$) indicates that the UCB-over-experts overhead reduces sample efficiency when a single algorithm clearly dominates.

    \item \textbf{Offline evaluation.} Our real-data experiments use offline replay of interaction logs, which introduces distributional shift: the algorithm's selected arm may differ from the arm that was actually presented, making counterfactual evaluation approximate.

    \item \textbf{Scalability.} The Whittle-advantage algorithms require solving single-arm MDPs via value iteration, which becomes expensive for fine-grained state discretizations.
    At $K{=}100$ this adds non-trivial overhead.

    \item \textbf{Fixed hyperparameters.} We use a single exploration constant $c = \sqrt{2}$ for all UCB-family algorithms.
    Per-algorithm tuning could improve individual performance but would complicate fair comparison.

    \item \textbf{Single-arm activation.} Our RMAB formulation assumes exactly one arm is activated per timestep.
    Real educational settings may allow multiple questions per session, requiring combinatorial extensions.
\end{itemize}
