\documentclass{article}

% --- Submission mode (choose ONE) ---
\usepackage[preprint]{neurips_2025}
% \usepackage{neurips_2025}          % anonymous submission
% \usepackage[final]{neurips_2025}   % camera-ready

% === Core packages ===
\usepackage[utf8]{inputenc}    % allow utf-8 input
\usepackage[T1]{fontenc}       % use 8-bit T1 fonts
\usepackage{hyperref}          % hyperlinks
\usepackage{url}               % simple URL typesetting
\usepackage{booktabs}          % professional-quality tables
\usepackage{amsfonts}          % blackboard math symbols
\usepackage{amsmath}           % math environments
\usepackage{amssymb}           % math symbols
\usepackage{nicefrac}          % compact symbols for 1/2, etc.
\usepackage{microtype}         % microtypography
\usepackage{xcolor}            % colors
\usepackage{caption}           % caption control

% === Figure and table packages ===
\usepackage{graphicx}
\graphicspath{{figs/}}
\usepackage{subcaption}        % subfigures
\usepackage{wrapfig}           % wrapped figures/tables
\usepackage{multirow}          % multi-row table cells
\usepackage{makecell}          % line breaks in table cells
\usepackage{colortbl}          % colored table rows

% === Cross-referencing ===
\usepackage[capitalize,noabbrev]{cleveref}

% === Checkmarks for comparison tables ===
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

% === Highlight color for "ours" row ===
\definecolor{blond}{rgb}{0.98, 0.94, 0.75}

% === Method name macro ===
\newcommand{\ours}{UCB-Quiz\xspace}
\usepackage{xspace}

% === Stat macros ===
\newcommand{\numstudents}{100\xspace}
\newcommand{\numcategories}{6\xspace}
\newcommand{\numquestions}{10{,}000\xspace}
\newcommand{\peakweakest}{60.7\%\xspace}
\newcommand{\peakoverall}{3.9\%\xspace}

% === Author block ===
\title{Adaptive Question Selection via Upper Confidence Bound\\for Efficient Knowledge Acquisition: A Simulation Study}

\author{%
\textbf{Yousef Radwan}\\
Independent Researcher
}

\begin{document}
\maketitle

\input{sec/0_abstract}
\input{sec/1_intro}
\input{sec/2_related}
\input{sec/3_method}
\input{sec/4_experiments}
\input{sec/5_conclusion}

\newpage
{\small
\bibliographystyle{unsrt}
\bibliography{refs}
}

\newpage
\input{sec/supplementary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section*{NeurIPS Paper Checklist}

\begin{enumerate}

\item {\bf Claims}
    \item[] Question: Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \item[] Answer: \answerYes{}
    \item[] Justification: All claims are supported by simulation results presented in \cref{sec:experiments}. We clearly state the simulation-based nature of the study.

\item {\bf Limitations}
    \item[] Question: Does the paper discuss the limitations of the work performed by the authors?
    \item[] Answer: \answerYes{}
    \item[] Justification: Limitations are discussed in \cref{sec:conclusion} and Section~S3 of the supplementary material, including the use of simulated (not real) students and simplified learning dynamics.

\item {\bf Theory Assumptions and Proofs}
    \item[] Question: For each theoretical result, does the paper provide the full set of assumptions and a complete (correct) proof?
    \item[] Answer: \answerNA{}
    \item[] Justification: This is an empirical simulation study. We reference the theoretical foundations of UCB1 from prior work~\cite{auer2002finite}.

\item {\bf Experimental Result Reproducibility}
    \item[] Question: Does the paper fully disclose all the information needed to reproduce the main experimental results of the paper to the extent that it affects the main claims and/or conclusions of the paper?
    \item[] Answer: \answerYes{}
    \item[] Justification: All simulation parameters are reported in \cref{sec:method}, the random seed is fixed at 42, and we provide complete source code.

\item {\bf Open Access to Data and Code}
    \item[] Question: Does the paper provide open access to the data and code, with sufficient instructions to faithfully reproduce the main experimental results?
    \item[] Answer: \answerYes{}
    \item[] Justification: The complete source code, experiment runner, and result data will be released on GitHub.

\item {\bf Experimental Setting/Details}
    \item[] Question: Does the paper specify all the training and test details necessary to understand the results?
    \item[] Answer: \answerYes{}
    \item[] Justification: All parameters (learning rate, decay rate, exploration constant, population size, number of questions) are specified in \cref{tab:config} and \cref{sec:method}.

\item {\bf Experiment Statistical Significance}
    \item[] Question: Does the paper report error bars suitably and correctly defined or other appropriate information about the statistical significance of the experiments?
    \item[] Answer: \answerYes{}
    \item[] Justification: We report standard deviations across the student population and analyze results across three independent runs with different seeds.

\item {\bf Experiments Compute Resources}
    \item[] Question: For each experiment, does the paper provide sufficient information on the computer resources needed to reproduce the experiment?
    \item[] Answer: \answerYes{}
    \item[] Justification: Computational requirements are minimal (standard CPU, under 1 minute per run) and are described in the supplementary material.

\item {\bf Code of Ethics}
    \item[] Question: Does the research conducted in the paper conform, in every respect, with the NeurIPS Code of Ethics?
    \item[] Answer: \answerYes{}
    \item[] Justification: This simulation study uses no real student data and raises no ethical concerns.

\item {\bf Broader Impacts}
    \item[] Question: Does the paper discuss both potential positive societal impacts and negative societal impacts of the work?
    \item[] Answer: \answerYes{}
    \item[] Justification: We discuss the potential for more efficient personalized learning and note the need for validation with real students before deployment.

\item {\bf Safeguards}
    \item[] Question: Does the paper describe safeguards that have been put in place for responsible release of data or models?
    \item[] Answer: \answerNA{}
    \item[] Justification: The released code is a simulation framework with no potential for misuse.

\item {\bf Licenses for Existing Assets}
    \item[] Question: Are the creators or original owners of assets used in the paper properly credited?
    \item[] Answer: \answerYes{}
    \item[] Justification: We cite the Gymnasium framework~\cite{towers2024gymnasium} and all algorithmic foundations.

\item {\bf New Assets}
    \item[] Question: Are new assets introduced in the paper well documented and is the documentation provided alongside the assets?
    \item[] Answer: \answerYes{}
    \item[] Justification: The simulation framework includes documentation, a README, and CLI help.

\item {\bf Crowdsourcing and Research with Human Subjects}
    \item[] Question: For crowdsourcing experiments and research with human subjects, does the paper include the full text of instructions given to participants?
    \item[] Answer: \answerNA{}
    \item[] Justification: No human subjects were involved; all students are simulated.

\item {\bf Declaration of LLM usage}
    \item[] Question: Does the paper describe the usage of LLMs in the research process?
    \item[] Answer: \answerYes{}
    \item[] Justification: An LLM (Claude) was used to assist with paper writing and code documentation. No LLMs were used in the experimental methodology or data generation.

\end{enumerate}

\end{document}
